---
title: "C3 - Data transformation"
output:
  html_document: default
date: "2024-02-13"
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(results='hide')
```

```{r include=FALSE}
library(nycflights13)
library(tidyverse)
```

## C3.2 Rows

```{r}
# A shorter way to select flights that departed in January or February, %in% means "is in" or "is one of"
flights |> # used on row, not change the ordinal data
  filter(month %in% c(1, 2))

flights |> # just arrange the data, not change the ordinal data
  arrange(year, month, day, dep_time)

# Remove duplicate rows, if any
flights |>
  distinct()

# Find all unique origin and destination pairs
flights |>
  distinct(origin, dest)
# to keep all columns, instead return just selected columns, `.keep_all` default is `FALSE`
flights |>
  distinct(origin, dest, .keep_all = TRUE)
# Note that it will just return the first unique one.

flights |>
  count(origin, dest, sort = TRUE)
# The `count` is recommend by the books to count the unique row.
# But even just using `distinct`, the result will also show the tibble 's size
```

### Exercise 1

1.  arrival delay for tow or more hours

```{r}
flights |>
  filter(arr_delay >= 120)
```

2.  flew to Houston

```{r}
flights |>
  filter(dest %in% c("IAH", "HOU"))
```

3.  were operated by United, American, or Delta

```{r}
# use `airlines` to check the carrier
flights |>
  filter(carrier %in% c("UA", "AA", "DL"))
```

4.  departed in summer (July, August, and September)

```{r}
flights |>
  filter(
    month %in% c(7:9),
  )
```

5.  arrived more than two hours late, but didn't leave late

```{r}
flights |>
  filter(arr_delay >= 120, dep_delay <= 0)
```

6.  were delayed by at least an hour, but made up over 30 minutes in flight

```{r}
flights |>
  filter(arr_delay >= 60, air_time > 30)
```

### Exercise 2

Find the most departure delayed flights

```{r}
flights |>
  arrange(desc(dep_delay)) |>
  slice_head(n = 1)
```

Find the earliest departure flights

```{r}
flights |>
  arrange(dep_time) |>
  slice_head(n = 5)
```

### Exercise 3

Sort flights with time in the air

```{r}
flights |>
  arrange(arr_time - dep_time)
```

### Exercise 4

Was there a flight on every day

```{r}
flights |>
  distinct(year, month, day) |>
  count() # yes, the result is 365
```

### Exercise 5

the farthest flights and the least flights

```{r}
flights |>
  arrange(distance) |>
  slice_head(n = 1)

flights |>
  arrange(desc(distance)) |>
  slice_head(n = 1)
```

### Exercise 6

If want to use `filter` and `arrange` together, the order is matter. Depending on specific analysis. Generally, if you're filtering a large data-set and only need a subset of rows to be sorted, it's better to filter first to reduce the amount of data being sorted. However, if you need to sort the entire data-set and then filter based on specific criteria, arranging first might be more appropriate.

## C3.3 Columns

```{r}
flights |>
  mutate(
    gain = dep_delay - arr_delay,
    speed = distance / air_time * 60,
    .after = day
  )
flights |>
  mutate(
    gain = dep_delay - arr_delay,
    hours = air_time / 60,
    gain_per_hour = gain / hours,
    .keep = "used"
  )

# There is many ways in `select`, should check details in the help page.
flights |>
  select(!year:day)
flights |>
  select(where(is.character))
flights |>
  select(tail_num = tailnum) # one way to rename the column

flights |>
  rename(tail_num = tailnum) # another way to rename the column
#  janitor::clean_names() is a good way to clean the column names
```

### Exercise 1

1.  Compare `dep_time`, `sched_dep_time`, and `dep_delay`. How would you expect those three numbers to be related?

```{r}
# use`dep_time` - `sched_dep_time` to calculate the delay
# Compare the `delay` with `dep_delay`
flights <- flights |>
  mutate(
    dep_time_diff = dep_time - sched_dep_time,
    comparison = dep_time_diff == dep_delay
  )
```

Most of them are same, but their difference not same, they're all multiples of ten.

```{r}
summary(flights$comparison) 

flights |>
  select(comparison, dep_time_diff, dep_delay) |>
  filter(comparison == FALSE) |>
  head(10)

flights |>
  select(comparison, dep_time_diff, dep_delay) |>
  filter(comparison == FALSE) |>
  tail(10)

# delete the created `dep_time_diff` and `comparison` columns
flights <- flights |>
  select(-dep_time_diff, -comparison)
```

### Exercise 2

2.  Brainstorm as many ways as possible to select `dep_time`, `dep_delay`, `arr_time`, and `arr_delay` from `flights`.

```{r}
flights |>
  select(dep_time, dep_delay, arr_time, arr_delay)

flights |>
  select(starts_with(c("dep", "arr")))

flights |>
  select(matches("^(dep|arr).*?(time|delay)$"))

flights |>
  select(all_of(c("dep_time", "dep_delay", "arr_time", "arr_delay")))

flights |>
  select(contains("dep"), contains("arr")) |>
  select(-starts_with("sched")) |>
  select(-carrier)
```

### Exercise 3

3.  What happens if you specify the name of the same variable multiple times in a [`select()`](https://dplyr.tidyverse.org/reference/select.html) call?

```{r}
flights |>
  select(dep_time, dep_time, dep_time) # just return one column
```

### Exercise 4

4.  What does the [`any_of()`](https://tidyselect.r-lib.org/reference/all_of.html) function do? Why might it be helpful in conjunction with this vector?

```{r}
variables <- c("year", "month", "day", "dep_delay", "arr_delay")
flights |>
  select(any_of(variables))
```

What is interesting is that it's difference between `any_of` and `all_of`. `any_of` will return the columns that are in the vector, if some columns are not in the vector, it will not return them. But `all_of` will throw an error if some columns are not in the vector. As said in document, `any_of` is useful when negative selection.

### Exercise 5

5.  Does the result of running the following code surprise you? How do the select helpers deal with upper and lower case by default? How can you change that default?

```{r}
flights |> select(contains("TIME")) # Capital non-sensitive
flights |> select(contains("TIME", ignore.case = FALSE)) # Capital sensitive
```

### Exercise 6

6.  Rename `air_time` to `air_time_min` to indicate units of measurement and move it to the beginning of the data frame.

```{r}
flights |>
  rename(air_time_min = air_time) |>
  relocate(air_time_min)
```

### Exercise 7

7.  Why doesn’t the following work, and what does the error mean?

```{r}
flights |> 
  select(tailnum) |> 
  arrange(arr_delay)
#> Error in `arrange()`:
#> ℹ In argument: `..1 = arr_delay`.
#> Caused by error:
#> ! object 'arr_delay' not found
```

Because the order of the `arrange` and `select` is wrong, the `arr_delay` is not in the data frame after the `select` function. So, the `arrange` function can't find the `arr_delay` column.

```{r}
flights |> 
  arrange(arr_delay) |> 
  select(tailnum) # works well
```

## C3.5 Groups

```{r}
flights |>
  group_by(month) |>
  summarize(
    avg_delay = mean(dep_delay, na.rm = TRUE),
    n = n()
  )

flights |>
  group_by(dest) |>
  slice_max(arr_delay, n = 1) |> # with_ties = TRUE to drop the same value rows
  relocate(dest)

daily <- flights |>  
  group_by(year, month, day)
daily |> 
  summarize(n = n(),
            .groups = "keep")

daily |>
  ungroup() |>
  summarize(
    avg_delay = mean(dep_delay, na.rm = TRUE), 
    flights = n()
  )

flights |> 
  summarize(
    delay = mean(dep_delay, na.rm = TRUE), 
    n = n(),
    .by = c(origin, dest) # per-operation grouping
  )
```

### Exercise

1.  Which carrier has the worst average delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about `flights |> group_by(carrier, dest) |> summarize(n())`)

F9 has the longest average delays, but when consider the counts, EV will be the worst.

```{r}
flights |>
  group_by(carrier) |>
  summarize(
    avg_delay = mean(dep_delay, na.rm = TRUE),
    n = n()
  ) |>
  arrange(desc(avg_delay))
```

In my opinion, the main effect of delay is bad carriers.

```{r}
flights |>
  group_by(dest) |>
  summarize(
    avg_delay = mean(dep_delay, na.rm = TRUE),
    n = n()
  ) |>
  arrange(desc(avg_delay))

flights |>
  summarize(
    avg_delay = mean(dep_delay, na.rm = TRUE),
    n = n(),
    .by = c(carrier, dest) # instead of `group_by`, use `.by` to suppress warning
  ) |>
  arrange(desc(avg_delay))
```

2.  Find the flights that are most delayed upon departure from each destination.

```{r}
flights |>
  group_by(dest) |>
  filter(dep_delay > 0) |> # drop the negative delay
  summarize(
    n = n()
  ) |>
  arrange(desc(n)) |>
  head(3) # LAX, ORD, ATL
```

3.  How do delays vary over the course of the day. Illustrate your answer with a plot.

```{r}
flights |>
  group_by(hour = dep_time %/% 100) |>
  summarize(
    avg_delay = mean(dep_delay, na.rm = TRUE),
    n = n()
  ) |>
  filter(!is.na(hour)) |> # if not, plot will raise an error
  ggplot() +
  geom_col(aes(x = hour, y = avg_delay), na.rm = TRUE) + 
  scale_x_continuous(breaks = 0:23) +
  labs(
    title = "Average delay over the course of the day",
    x = "Hour", y = "Average delay (minutes)"
  ) +
  theme_minimal()
```

4.  What happens if you supply a negative `n` to [`slice_min()`](https://dplyr.tidyverse.org/reference/slice.html) and friends?

```{r}
flights |>
  group_by(dest) |>
  slice_min(arr_delay, n = -1) |> # return all rows
  relocate(dest)
```

5.  Explain what [`count()`](https://dplyr.tidyverse.org/reference/count.html) does in terms of the dplyr verbs you just learned. What does the `sort` argument to [`count()`](https://dplyr.tidyverse.org/reference/count.html) do?

The codes below is return totally the same result.

```{r}
flights |>
  count(dest, sort = TRUE)

flights |>
  group_by(dest) |>
  summarize(n = n()) |>
  arrange(desc(n))
```

6.  Suppose we have the following tiny data frame:

```{r}
df <- tibble(
  x = 1:5,
  y = c("a", "b", "a", "a", "b"),
  z = c("K", "K", "L", "L", "K")
)
```

a.  Write down what you think the output will look like, then check if you were correct, and describe what [`group_by()`](https://dplyr.tidyverse.org/reference/group_by.html) does.

Group by `y`, return grouped data.

```{r}
df |> group_by(y)
```

b.  Write down what you think the output will look like, then check if you were correct, and describe what [`arrange()`](https://dplyr.tidyverse.org/reference/arrange.html) does. Also comment on how it’s different from the [`group_by()`](https://dplyr.tidyverse.org/reference/group_by.html) in part (a)?

The difference between before is the return object is not grouped, just arranged.

```{r}         
df |> arrange(y)
```

c.  Write down what you think the output will look like, then check if you were correct, and describe what the pipeline does.

Group data by `y`, then calculate the mean of `x` in each group (a, b).

```{r}         
df |>
  group_by(y) |>
  summarize(mean_x = mean(x))
```

d.  Write down what you think the output will look like, then check if you were correct, and describe what the pipeline does. Then, comment on what the message says.

Group data by `y` and `z`, then calculate the mean of `x` in each group (a, b) and (K, L) union. The message is to remind that the output is first grouped by `y`.

```{r}         
df |>
  group_by(y, z) |>
  summarize(mean_x = mean(x))

df |>
  group_by(y, z) |>
  summarize(mean_x = mean(x), .groups = "drop") # to return un-grouped data
```

e.  Write down what you think the output will look like, then check if you were correct, and describe what the pipeline does. How is the output different from the one in part (d).

```{r}         
df |>
  group_by(y, z) |>
  summarize(mean_x = mean(x), .groups = "drop")
```

f.  Write down what you think the outputs will look like, then check if you were correct, and describe what each pipeline does. How are the outputs of the two pipelines different?

The first one's result is grouped by `y` with warning, and has a new col naming `mean_x` from `summarize()`. 

The second one's result is grouped by `y` and `z`without warning, and has a new col naming `mean_x` created from `mutate()`.

```{r}         
df |>
  group_by(y, z) |>
  summarize(mean_x = mean(x))

df |>
  group_by(y, z) |>
  mutate(mean_x = mean(x))
```
## C3.6 Case Study

```{r}
batters <- Lahman::Batting |> 
  group_by(playerID) |> 
  summarize(
    performance = sum(H, na.rm = TRUE) / sum(AB, na.rm = TRUE),
    n = sum(AB, na.rm = TRUE)
  )
batters
```
Decrease variance and positive linear relationship.

```{r}
batters |> 
  filter(n > 100) |> 
  ggplot(aes(x = n, y = performance)) +
  geom_point(alpha = 1 / 10) + 
  geom_smooth(se = FALSE)
```
