---
title: "C5"
output: html_notebook
---

```{r}
knitr::opts_chunk$set(results='hide')

library(nycflights13)
library(tidyverse)
```

## C5.2 Tidy data

### Exercise

1.  For each of the sample tables, describe what each observation and each column represents.

This is easy to do by checking the data.

2.  Sketch out the process you’d use to calculate the `rate` for `table2` and `table3`. You will need to perform four operations:

    a.  Extract the number of TB cases per country per year.

    b.  Extract the matching population per country per year.

    c.  Divide cases by population, and multiply by 10000.

    d.  Store back in the appropriate place.

You haven’t yet learned all the functions you’d need to actually perform these operations, but you should still be able to think through the transformations you’d need.

```{r}
cases_count <- table2 |>
  filter(type == "cases") |>
  select(count)

population_count <- table2 |>
  filter(type == "population") |>
  select(count)

rate <- cases_count / population_count * 10000
rate
```
For `table3`, should transform the third column from string to equation and then calculate the rate. 


