---
title: "R Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(results = "hide")

library(nycflights13)
library(tidyverse)
library(janitor)
```

## C9.2 Aesthetic mappings

### Exercise

1.  create a scatterplot of `hwy` vs. `displ` where the points are pink filled in triangles.

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(shape = 17, color = "pink", fill = "pink")
```

2.  Why did the following code not result in a plot with blue points?

color = "blue" should be placed as an argument of `geom_point`, not inside `aes`.

```{r}
ggplot(mpg) +
  geom_point(aes(x = displ, y = hwy), color = "blue")
```

3.  What does the `stroke` aesthetic do? What shapes does it work with? (Hint: use `?geom_point`)

`stroke` is used to set the line's width.

4.  What happens if you map an aesthetic to something other than a variable name, like `aes(color = displ < 5)`? Note, you’ll also need to specify `x` and `y`.

It use a logical condition to colored the point in blue. Note that, should not set `color` as an argument of `geom_point` in this case. `geom_point(color="blue", aes(color = displ < 5))` will set all points to blue.

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(aes(color = displ < 5))
```

## C9.3 Geometric objects

The below chat's red points are created two layers, red points and red open circles.

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  geom_point(
    data = mpg |> filter(class == "2seater"),
    color = "red"
  ) +
  geom_point(
    data = mpg |> filter(class == "2seater"),
    shape = "circle open", size = 3, color = "red"
  )
```

```{r}
library(ggridges)

ggplot(mpg, aes(x = hwy, y = drv, fill = drv, color = drv)) +
  geom_density_ridges(alpha = 0.5, show.legend = FALSE)
```

### Exercise

1.  What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_line()

ggplot(mpg, aes(x = class, y = hwy)) +
  geom_boxplot()

ggplot(mpg, aes(x = hwy)) +
  geom_histogram()
```

2.  Earlier in this chapter we `used show.legend` without explaining it:

As shown below, it used to hide the legend of the geom.

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_smooth(aes(color = drv), show.legend = FALSE)

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_smooth(aes(color = drv))
```

3.  What does the `se` argument to `geom_smooth()` do?

It used to control whether show the confidence level.

4.  Recreate the R code necessary to generate the following graphs. Note that wherever a categorical variable is used in the plot, it’s `drv`.

```{r}
# This is improved by ChatGPT
library(patchwork)

base_plot <- function() {
  ggplot(mpg, aes(x = displ, y = hwy)) +
    geom_point(color = "black") +
    theme_minimal()
}

p1 <- base_plot() +
  geom_smooth(color = "blue", se = FALSE)

p2 <- base_plot() +
  geom_smooth(aes(group = drv), color = "blue", se = FALSE)

p3 <- base_plot() +
  geom_point(aes(color = drv)) + geom_smooth(aes(group = drv), se = FALSE)

p4 <- base_plot() +
  geom_point(aes(color = drv)) + geom_smooth(color = "blue", se = FALSE)

p5 <- base_plot() +
  geom_point(aes(color = drv)) + geom_smooth(color = "blue", se = FALSE, aes(linetype = drv))

p6 <- base_plot() +
  geom_point(aes(color = drv)) + geom_point(shape = "circle open", size = 3, color = "white")

# Combine plots using patchwork
(p1 | p2) / (p3 | p4) / (p5 | p6)
```

## C9.4 Facets

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  # for categorical variables, and stay y-axis free
  facet_grid(drv ~ cyl, scales = "free_y")
```

### Exercise

1.  What happens if you facet on a continuous variable?

As shown below, `cty` is a continuous variable, if facet on it, the plot will be split by each value of `cty` times each value of `drv`(combination of their values).

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  facet_grid(drv ~ cty, scales = "free_y")
```

2.  What do the empty cells in the plot above with `facet_grid(drv ~ cyl)` mean? Run the following code. How do they relate to the resulting plot?

The empty cell indicates that there is no data in the corresponding/particular combination of `drv` and `cyl`.

```{r}
ggplot(mpg) +
  geom_point(aes(x = drv, y = cyl))
```

3.  What plots does the following code make? What does `.` do?

```{r}
ggplot(mpg) +
  geom_point(aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(mpg) +
  geom_point(aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

`.` represents all variables in dataset. For the fist one, `drv ~ .` specifies that should facet by `drv` in rows(on y-axis), and `.` means columns are all other variables in dataset. It will show all data in dataset with specific level of `drv` as demo codes to show all data with `drv == 4` below, it as same as the first row in above first chat.

```{r}
ggplot(
  mpg |> filter(drv == 4) |> select(displ, hwy),
  aes(x = displ, y = hwy)
) +
  geom_point()
```

4.  Take the first faceted plot in this section:

```{r}
ggplot(mpg) +
  geom_point(aes(x = displ, y = hwy)) +
  facet_wrap(~class, nrow = 2)

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(aes(color = class))
```

What are the advantages to using faceting instead of the color aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

The advantage of using faceting is that it can show more details of the data, and the disadvantage is that it will make the plot more complex. If you have a larger dataset, the balance will change, and the advantage of using faceting will be more obvious, especially when you have more levels of specific categorical variables.

5.  Read `?facet_wrap`. What does `nrow` do? What does `ncol` do? What other options control the layout of the individual panels? Why doesn’t `facet_grid()` have `nrow` and `ncol` arguments?

`nrow` and `ncol` are used to control the number of rows and columns of the facet. Other options control the layout of the individual panels are `scales`, `shrink`, `labeller`, `as.table`, `switch`.

`facet_grid()` doesn't have `nrow` and `ncol` arguments because it can specify the layout of the individual panels by `rows` and `columns` directly, and usually it used as below to control what the combination of categorical variables, not like `facet_wrap()` to control the layout of the individual panels.

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  facet_grid(rows = vars(drv), cols = vars(cyl))
```

6. Which of the following plots makes it easier to compare engine size (`displ`) across cars with different drive trains? What does this say about when to place a faceting variable across rows or columns?

The difference between below plots is using level of `drv` as rows or columns. To compare `displ` with different level of `drv`, I think the second plot, using level of `drv` as columns corresponding using `displ` as x-axis, is more suitable as each columns represents a different `dev`, and you can compare the distribution of `displ` in different `drv` more easily. 

This suggests that when you want to compare a continuous variable across different levels of a categorical variable, you should place the faceting variable across columns rather than rows. That allows side-by-side comparison of the distribution within each category of the faceting variable.

```{r}
ggplot(mpg, aes(x = displ)) + 
  geom_histogram() + 
  facet_grid(drv ~ .)

ggplot(mpg, aes(x = displ)) + 
  geom_histogram() +
  facet_grid(. ~ drv)
```

7. Recreate the following plot using `facet_wrap()` instead of `facet_grid()`. How do the positions of the facet labels change?

As seen below, the first using `facet_grid()`, and it will treat different level of `drv` as rows, and the second one using `facet_wrap()` using levels of `drv` as columns.

```{r}
ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(mpg) +
  geom_point(aes(x = displ, y = hwy)) +
  facet_wrap(~drv)
  # facet_wrap(~drv, ncol = 1)
```
